---
title: 'P4: Explore and Sumarize Data'
author: "David Gonzalez Cagigas"
date: "7 Februar 2017"
output: html_document
---
========================================================

```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE)
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.
#
#install.packages("ggplot2", dependencies = T) 
#install.packages("knitr", dependencies = T)
#install.packages("dplyr", dependencies = T)
#install.packages('gridExtra') 
#install.packages("GGally")
source("https://raw.githubusercontent.com/briatte/ggcorr/master/ggcorr.R")

library(gridExtra) 
library(ggplot2)
library(knitr)
library(dplyr)

```

First of all, we will load the data set Red Wine into pf, and see some basic
information.

```{r, Load_the_Data}

pf <- read.csv('wineQualityReds.csv')# Load the Data
pf$X <- NULL
names(pf)
str(pf)
summary(pf)

```

We can appreciate the variable names in the dataset, number of obervations, 
number and type of variables and other basic statistic information about the 
dataset. I have omitted the first variable as it is just the index.

If we dive deeper, we see that there are 6 qualities (from 3 to 8), the alcohol
degrees is between 8.40 and 14.90 degrees. Density vary from 0.9901 to 1.0037. 
And 10 more variables. It will be intersting to see the pH concentration,
alcohol or density depending on the quelity level.

# Univariate Plots Section
> In this section, you should perform some preliminary exploration of
your dataset. Run some summaries of the data and create univariate plots to
understand the structure of the individual variables in your dataset. Don't
forget to add a comment after each plot or closely-related group of plots!
There should be multiple code chunks and text sections; the first one below is
just to help you get started.

We will start showing one of the most important features, which is the quality
of the wine. With this plot we weill have an idea about what is the most common
quality in wines.

```{r}

by(pf$density, pf$quality, median)
by(pf$alcohol, pf$quality, median)
by(pf$pH, pf$quality, median)

qplot(factor(quality), data=pf, geom="bar")

```

As you can see, density and alcohol do not have a dependency on que quality of 
the wines, however quality seems to improve when Ph mean decrease. It is useful 
to know the number of observation for each quality.

Histograms

```{r}
grid.arrange( ggplot(aes(x=alcohol),data=pf) +
  geom_histogram( binwidth=0.5) + 
  ggtitle("Histogram of Alcohol (%)"), ggplot(aes(x=1, y=alcohol),data=pf) +
  geom_boxplot( ), nrow =1)
summary(pf$alcohol)
qplot(x = alcohol, data = pf, binwidth = 0.5, geom = "freqpoly")

```

Alcohol distribution is a right skewed, concentrated in 7.5. We can see a few 
outliers. It seems to be more common the low alcohol wines than the high alcohol
wine. As you can see the most common wines have unde 10 grades.

```{r}

grid.arrange( ggplot(aes(x=total.sulfur.dioxide),data=pf) +
  geom_histogram( binwidth=10) + 
  ggtitle("Histogram of total sulfur dioxide (mg/dm^3)"), 
  ggplot(aes(x=1, y=total.sulfur.dioxide),data=pf) +
  geom_boxplot( ), nrow =1)
summary(pf$total.sulfur.dioxide)
qplot(x = total.sulfur.dioxide, data = pf, binwidth = 10, geom = "freqpoly")

```

Sulfur Dioxide distribution is a right skewed, concentrated in 20. We can see
multiples outliers, specially 2. With the Sulfur Dioxide happens the same as 
with alcohol. Most of wines have a low quantity of sulfur dioxide, the peak is
at 40g/dm3.


```{r}

grid.arrange( ggplot(aes(x=fixed.acidity),data=pf) +
  geom_histogram( binwidth=1) + 
  ggtitle("Histogram of fixed acidity (g/dm^3)"),
  ggplot(aes(x=1, y=fixed.acidity),data=pf) +
  geom_boxplot( ), nrow =1)
summary(pf$fixed.acidity)
qplot(x = fixed.acidity, data = pf, binwidth = 1, geom = "freqpoly")

```

We can see the fixed acidity is a right skewed graphic concentred in 7. We can
see multiple outliers. Fixed acidity has a similar distribution as the other 
features with its peak at 7g/dm3.

```{r}

grid.arrange( ggplot(aes(x=volatile.acidity),data=pf) +
  geom_histogram( binwidth=0.05) + 
  ggtitle("Histogram of volatile acidity (g/dm^3)"), 
  ggplot(aes(x=1, y=volatile.acidity),data=pf) +
  geom_boxplot( ), nrow =1)
summary(pf$volatile.acidity)
qplot(x = volatile.acidity, data = pf, binwidth = 0.05, geom = "freqpoly")

```

The distribution of volatile acidity is a bimodal right skewed, concentred in 
0.35 and 0.6. There are a few outliers. Volatile acidity has a similar 
distribution as the other features, however volatile acidity has a broad peak
between 0.4 and 0.7 g/dm3.

```{r}

grid.arrange( ggplot(aes(x=citric.acid),data=pf) +
  geom_histogram( binwidth=0.05) +
  ggtitle("Histogram of citric acid (g/dm^3)"),
  ggplot(aes(x=1, y=citric.acid),data=pf) +
  geom_boxplot( ), nrow =1)
summary(pf$citric.acid)
qplot(x = citric.acid, data = pf, binwidth = 0.05, geom = "freqpoly")

```

Citrid Acid is a not normal distribution. Only an outlier. It doesn't seem to be
a common citrid acid for the wines in the data set.

```{r}

grid.arrange( ggplot(aes(x=residual.sugar),data=pf) + 
              coord_cartesian(xlim = c(1,5)) +
  geom_histogram( binwidth=0.5) + 
  ggtitle("Histogram of residual sugar (g/dm^3)"),
  ggplot(aes(x=1, y=residual.sugar),data=pf) +
  geom_boxplot( ) , nrow =1)
summary(pf$residual.sugar)
qplot(x = residual.sugar, data = pf, binwidth = 0.5, geom = "freqpoly") +
  coord_cartesian(xlim = c(1,5))


```

Several outliers. As most of the features, it is a right skewed distribution 
concentrated in 2 g/dm3. 

```{r}

grid.arrange( ggplot(aes(x=chlorides),data=pf) + 
                coord_cartesian(xlim = c(0,0.2)) +
  geom_histogram( binwidth=0.01) +
  ggtitle("Histogram of chlorides (g/dm^3)"),
  ggplot(aes(x=1, y=chlorides),data=pf) +
  geom_boxplot( ), nrow =1)
summary(pf$chlorides)
qplot(x = chlorides, data = pf, binwidth = 0.01, geom = "freqpoly") +
  coord_cartesian(xlim = c(0,0.2))


```

Chloricles is right skewed distribution as well, concentrated in 0.08. Several
outliers. Chloricles has a similar distribution as the other 
features with its peak at 0.08 g/dm3.

```{r}

grid.arrange( ggplot(aes(x=free.sulfur.dioxide),data=pf) +
  geom_histogram( binwidth=5) +
  ggtitle("Histogram of free sulfur dioxide (mg/dm^3)"),
  ggplot(aes(x=1, y=free.sulfur.dioxide),data=pf) +
  geom_boxplot( ), nrow =1)
summary(pf$free.sulfur.dioxide)
qplot(x = free.sulfur.dioxide, data = pf, binwidth = 5, geom = "freqpoly")


```

Free Sulfur Dioxide is a right skewed distribution concentrated in 6. We can see
a few outliers. Free Sulfur Dioxide has a similar distribution as the other 
features with its peak at 10 mg/dm3.

```{r}

grid.arrange( ggplot(aes(x=density),data=pf) +
  geom_histogram( binwidth=0.001) +
  ggtitle("Histogram of density (g/cm^3)"),
  ggplot(aes(x=1, y=density),data=pf) +
  geom_boxplot( ), nrow =1)
summary(pf$density)
qplot(x = density, data = pf, binwidth = 0.001, geom = "freqpoly")


```

We can see the density distribution is normal and concentrated in 0.997. We can
see a few outliers.

```{r}

grid.arrange( ggplot(aes(x=sulphates),data=pf) +
  coord_cartesian(xlim = c(0,1.5)) +
  geom_histogram( binwidth=0.1) +
  ggtitle("Histogram of sulphates (g/dm^3)"),
  ggplot(aes(x=1, y=sulphates),data=pf) +
  geom_boxplot( ), nrow =1)
summary(pf$sulphates)
qplot(x = sulphates, data = pf, binwidth = 0.1, geom = "freqpoly") +
  coord_cartesian(xlim = c(0,1.5))


```

Sulphates distribution is right skewed, concentrated in 0.6. Several outliers.
Free Sulfur Dioxide has a similar distribution as the other features with its 
peak at 0.6 g/dm3.

```{r}

grid.arrange( ggplot(aes(x=pH),data=pf) +
  geom_histogram( binwidth=0.1) +
  ggtitle("Histogram of pH"),
  ggplot(aes(x=1, y=pH),data=pf) +
  geom_boxplot( ), nrow =1)
summary(pf$pH)
qplot(x = pH, data = pf, binwidth = 0.1, geom = "freqpoly")



```

pH distribution is normal concentrated in 3.3. We can appreciate a few outliers.

We can see above a simple histogram for each feature included in the analysis.  
Theses histograms help you to have an idea about the dataset. 

```{r}

qplot(x = density, data = pf, binwidth = 0.0001)+ ##A line geometry
    facet_wrap(~quality)+
    geom_histogram()
qplot(x = alcohol, data = pf, binwidth = 0.01)+ ##A line geometry
    facet_wrap(~quality)+
    geom_histogram()
qplot(x = pH, data = pf, binwidth = 0.01)+ ##A line geometry
    facet_wrap(~quality)+
    geom_histogram()


```

We can appreciate that pH tend to have similar values between 3 and 3.5, 
specially in quality 5, 6 and 7 the other qualities can not be appriciated due
to lack of data. Density seems to have the same tendency, however alcohol
have a flater graphs, where you can see a tendency in quality 5 (about 11 
degrees), meanwhile quality 7 is between 9 and 13 or quality 6 is between
8 and 13.


# Univariate Analysis

### What is the structure of your dataset?
There are 1599 Red Wines analysed in 12 different characteristics 
(fixed.acidity, volatile.acidity, citric.acid, residual.sugar, chlorides, 
free.sulfur.dioxide, total.sulfur.dioxidedensity, pH, sulphates, alcohol and
quality)

### What is/are the main feature(s) of interest in your dataset?
The main feature is the quality. It is interesting to caompare any 
characteristic of the wine depending on the final quality.

### What other features in the dataset do you think will help support your 
investigation into your feature(s) of interest?
I use all the features of the wine to compare theis variation of them with the
quality or bewteen them.

### Did you create any new variables from existing variables in the dataset?

### Of the features you investigated, were there any unusual distributions? Did 
you perform any operations on the data to tidy, adjust, or change the form of 
the data? If so, why did you do this?
Since the data is cleaned, I didn't perfomr any cleaning process. I realized 
most of graphs are right skewed. And graphs with vlaues of the highest quality
do not have a defined shape. It might be because of the lack of samples.

# Bivariate Plots Section
> It might be interesting a lot of correlations. As we have 12 
variables we will try with several combinations.

We will start the bivariate section with a correlation analysis that can help
us to decide which features are more likely to have a relation.

```{r}

ggcorr(pf, nbreaks = 4, palette = "RdGy",
       label = TRUE, label_size = 3, label_color = "white")

```

As we can see on the above plot, the feature with most correlation are free.
sulfur.dioxide-total.sulfur.dioxide, fixed.acidity-citrid.acid, 
fixed.acidity-pH and density-dixed.acidity. 

Along this analysis we will try to focus on the features, but we will study
other features as well. We weill start with a scatter plot comparing the 
features with the highest grade of correlation.

```{r, Bivariate_Plots}

ggplot(aes(x= total.sulfur.dioxide, y = free.sulfur.dioxide) , data = pf) +
  geom_point(alpha = 0.3, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size=1) +
  scale_color_brewer(type='seq', guide=guide_legend(title='Quality'))

ggplot(aes(x= fixed.acidity, y = density) , data = pf) +
  geom_point(alpha = 0.3, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size=1) +
  scale_color_brewer(type='seq', guide=guide_legend(title='Quality'))

ggplot(aes(x= fixed.acidity, y = citric.acid) , data = pf) +
  geom_point(alpha = 0.3, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size=1) +
  scale_color_brewer(type='seq', guide=guide_legend(title='Quality'))

ggplot(aes(x= pH, y = fixed.acidity) , data = pf) +
  geom_point(alpha = 0.3, size = 1) +
  geom_smooth(method = "lm", se = FALSE,size=1) +
  scale_color_brewer(type='seq', guide=guide_legend(title='Quality')) 


```

I displayed four graphs showing the relation between 2 variables. We can see
easily a correlation between all the features compared. 

Total Sulfur Dioxide - Free Sulfur Dioxide 
The bigger Total Sulfur Dioxide is, the bigger Free Sulfur Dioxide is. It seems 
to be a proportional relation between these features.

Density - Fixed Acidity 
The bigger Density is, the bigger Fixed Acidity is. It seems to be a 
proportional relation between these features.

Citrid Acid - Fixed Acidity 
The bigger Density is, the bigger Fixed Acidity is. It seems to be a 
proportional relation between these features.

pH - Fixed Acidity 
The bigger Density is, the smaller Fixed Acidity is. It seems to be a 
proportional inverse relation between these features.

The result show what we have see with the correlation analysis, Fixed Acidity
and pH has a inverse proportional correlation meanwhile the other analysis has
a poportional correlation.


```{r}

quality_groups <- group_by(pf, quality)     # first groups data by age

pf.by_quality <- summarise(quality_groups,                 # then summarizes
          volatile.acidity_mean = mean(volatile.acidity),     
          fixed.acidity_mean = mean(fixed.acidity),     
          residual.sugar_mean = mean(residual.sugar),     
          chlorides_mean = mean(chlorides),    
          free.sulfur.dioxide_mean = mean(free.sulfur.dioxide),     
          total.sulfur.dioxide_mean = mean(total.sulfur.dioxide),     
          sulphates_mean = mean(sulphates),     
          density_mean = mean(density),     
          alcohol_mean = mean(alcohol),     
          pH_mean = mean(pH),     
          n = n()     # number of items in each group
          )

pf.by_quality <- arrange(pf.by_quality, quality) # and sort by age

pf.by_quality


```

We can see a mean table of each variable separated by quality level. We will see
easily correlation in some of them like chlorides_mean which decrease when the
quality decrease, or pH with a similar behaviour. We will display this graphs 
below.

```{r}

 ggplot(aes(factor(quality), alcohol), data = pf) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", geom = "point",
               color = "red", shape = 8, size = 4)

 ggplot(aes(factor(quality), volatile.acidity), data = pf) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", geom = "point",
               color = "red", shape = 8, size = 4)
 
 ggplot(aes(factor(quality), free.sulfur.dioxide), data = pf) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", geom = "point",
               color = "red", shape = 8, size = 4)
 
 ggplot(aes(factor(quality), sulphates), data = pf) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", geom = "point",
               color = "red", shape = 8, size = 4)
 
 ggplot(aes(factor(quality), pH), data = pf) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", geom = "point",
               color = "red", shape = 8, size = 4)
 
 ggplot(aes(factor(quality), density), data = pf) +
  geom_jitter( alpha = .3)  +
  geom_boxplot( alpha = .5,color = 'blue') +
  stat_summary(fun.y = "mean", geom = "point",
               color = "red", shape = 8, size = 4)

```

One of the most interesting variables is the quality. In the plot above, I 
compare this features with the other. We can see thanks to the box plot the 
distribution of every feature divided by quality.

```{r}
cor(pf$alcohol, pf$density)
cor(pf$residual.sugar, pf$density)
cor(pf$fixed.acidity, pf$density)
cor(pf$pH, pf$density)
cor(pf$alcohol, pf$volatile.acidity)
cor(pf$residual.sugar, pf$volatile.acidity)

```

There is barely a correlation between residual sugar and volatile acidity, as we
expected. However we can find a correlation of 0.668 between density and acidity
which means that the bigger density the wine will be more acid. Alcohol and 
density has a inverse correlation, it means the more alcohol the less dense the 
wine is.
It happen the same with alcohol and acidity to lesser extent.

```{r}

q1 <- ggplot(aes(x = volatile.acidity_mean, y = pH_mean), data = pf.by_quality) +
  geom_line() +
  geom_smooth()

q2 <- ggplot(aes(x = volatile.acidity_mean, y = free.sulfur.dioxide_mean), 
             data = pf.by_quality) + geom_line() + geom_smooth()

q3 <- ggplot(aes(x = density_mean, y = pH_mean), data = pf.by_quality) +
  geom_line() + geom_smooth()

q4 <- ggplot(aes(x = density_mean, y = fixed.acidity_mean), 
             data = pf.by_quality) + geom_line() + geom_smooth()

library(gridExtra)
grid.arrange(q1,q2,q3,q4,ncol = 1)

```

This three last graphs show the relation between pHacidity,sulfur 
dioxide-acidity and pH-density. We see relation betwen pH and acidity, it is
a left skewed graph while the second sample is a right skewed graph.


# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the 
investigation. How did the feature(s) of interest vary with other features in 
the dataset?
As we have said before, there are a lot of correlations between variables. One 
example is acidity and density. We got a correlation of 0.67. It seems that the
more dense is the wine the taste get more acid.


### Did you observe any interesting relationships between the other features 
(not the main feature(s) of interest)?
We can see many  correlations between most of variables. For example residual 
sugar and density, which seems logical as the more sugar more dense. Or pH and
density.

### What was the strongest relationship you found?
Density and acididity was the strongest one. 



# Multivariate Plots Section

```{r, Multivariate_Plots}

ggplot(aes(x = alcohol, y = citric.acid, color = factor(quality)), data = pf) +
      geom_point(alpha = 0.8, size = 2) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
      theme_dark() +
      scale_color_brewer(type='seq', guide=guide_legend(title='Quality'))

ggplot(pf, aes(x = sulphates, y = alcohol, colour=factor(quality))) + 
  geom_density2d(bins=2) + 
  scale_color_brewer() + 
        theme_dark() +
  geom_point(alpha=0.1) 

ggplot(aes(x = volatile.acidity, y = residual.sugar, color = factor(quality)),
       data = pf) + geom_point(alpha = 0.8, size = 2) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
        theme_dark() +
      scale_color_brewer(type='seq', guide=guide_legend(title='Quality'))

```

First graph we can appreciate a relation between quality and alcohol, as most of
white points are wines with low alcohol. However it is not easy to see a 
relation between density and pH. Density distribution does not depends on pH.

In the second graph seems clear, good quality winehas the biggest concentration 
of alcohol and sulphate.

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the 
investigation. Were there features that strengthened each other in terms of 
looking at your feature(s) of interest?
As we can see in last three graphs, alcohol, pH, sulphates and residual.sugar 
are important for wine quality however ther is not a perfect value for each
feature. The quality level depends on the sum up of every feature.
We can see high contained in alcohol and low residual.sugar with high quality
and oposite.

### Were there any interesting or surprising interactions between features?
Personally, I expected that all features tend to a perfect value. But as I said 
in last question, wine quelity is a summ of features.

------

# Final Plots and Summary
You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One

```{r, Plot_One}
qplot(factor(quality), data=pf, geom="bar", main = "Wines distribution")+
  xlab("Quality") + 
  ylab("Number of wines")

```

### Description One
This first graph represent the number of wines analyzed in each quality. We can
see most of wines have a quality of 5 or 6, meanwhile the qualities 3, 4 and 8
have barely wines.

### Plot Two

```{r, Plot_Two}
q1 <- ggplot(aes(x = quality, y = pH), data = pf) +
      geom_jitter( alpha = .1)  +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
      scale_color_brewer(type='seq', guide=guide_legend(title='Quality')) +
      xlab("Quality") + 
      ylab("pH") +
      ggtitle("Quality / pH")

q2 <- ggplot(aes(x = quality, y = free.sulfur.dioxide), data = pf) +
      geom_jitter( alpha = .1)  +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
      scale_color_brewer(type='seq', guide=guide_legend(title='Quality')) +
      xlab("Quality") + 
      ylab("Free Sulfur Dioxide (eq/L)") +
      ggtitle("Quality / SO2 Mean")


q3 <- ggplot(aes(x = quality, y = alcohol), data = pf) +
      geom_jitter( alpha = .1)  +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
      scale_color_brewer(type='seq', guide=guide_legend(title='Quality')) +
      xlab("Quality") + 
      ylab("Alcohol (%)") +
      ggtitle("Quality / Alcohol (%)")



q4 <- ggplot(aes(x = quality, y = fixed.acidity), data = pf) +
      geom_jitter( alpha = .1)  +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
      scale_color_brewer(type='seq', guide=guide_legend(title='Quality')) +
      xlab("Quality") + 
      ylab("Fixed Acidity (g/L)") +
      ggtitle("Quality / Fixed Acidity")

library(gridExtra)
grid.arrange(q1,q2,q3,q4,ncol = 2)

```



### Description Two
We can see 4 scatter plot comparing the quality with 4 different features,
pH, Free Sulfur Dioxide, Alcohol and Fixed Acidity. Here we can see the most
likely value for each level.

### Plot Three

```{r, Plot_Three}
ggplot(aes(x=pH, y=alcohol, color=factor(quality)), data = pf) + 
      geom_point(alpha = 0.8, size = 2) +
      geom_smooth(method = "lm", se = FALSE,size=1)  +
      scale_color_brewer(type='seq', guide=guide_legend(title='Quality')) +
      xlab("pH") + 
      theme_dark() +
      ylab("Alcohol Concentration (Degree)") +
      labs(color = "Quality") +
      ggtitle("pH, Alcohol, Quality and Density Graph Relation")



```

### Description Three
Here we can appreciate many relations. We can clearly see whiter colors are 
below while darker colors are above, it means quality and alcohol has a 
proportional relation. Ph does not seem to be related to any other parameter and
density seems to be lower with higher alcohol concentration.

------

# Reflection
The wines data set contains information on 1599 wines across twelve variables 
from around 2009. Being an example dataset I expected clear relations between 
its features, however barely you can find some relations. First thing surprised 
me was that most of wines measured had a quality of 5 or 6 while qualities as 3,
4 or 8 have very few samples. I guess it means that most of wines has a normal 
quality and just a few have a high or low quality.

The most cleare relations I saw was the alcohol concentration, it is related 
with quality and with density.

Another thing is every feature has a perfect value indicidually, as you can see
in the firsts graphs where I compare quality with every feature. However, when
you have a best quality wine, these values change. It is like a lower value in 
alcohol can compensate a higher value in sugar.
